Scaffolder - Software for Reproducible Genome Scaffolding.

### 2011-03-09 src/article.tex
Minor corrections

  \paragraph*{Methods:} The software outlined here, ``Scaffolder'', is
  implemented in the Ruby programming language and can be installed via the
  RubyGems software management system. Genome scaffolds are defined using YAML
  - a data format, which is both human and machine-readable. Command line
  binaries and extensive documentation are available.

  \paragraph*{Results:} This software allows a genome build to be defined in
  terms of the constituent sequences using a relatively simple syntax to
  define the scaffold. This syntax further allows  unknown regions to be
  defined, and add inserts to fill gaps in the scaffold. Defining the genome
  construction in a file makes the scaffolding process reproducible and easier
  to edit compared with FASTA nucleotide sequence.

@@

The advent of high throughput sequencing methods has led to a renewed interest
in algorithms to solve the problem of genome assembly \cite{pop2008,pop2009}.
Nonetheless, due to the complexity of merging large numbers of overlapping
reads genome assembly software may be unable to produce a complete sequence.
Instead the algorithm may generate several large assembled regions of sequence
(`contigs') composed of the many individual reads that represent assembly into
the longest possible sequences, given the data. Nonetheless, these contigs
represent a fragmented picture of the genome and require additional work to
join together into a complete sequence. \pb

@@

The software outlined here, ``Scaffolder'', aims to address these problems of
reproducibility by creating a file syntax and software framework for editing
a genome scaffold. Scaffolder uses a specific file format to define how contigs
are joined, additional sequences are inserted, and for the specification of
unknown regions. This syntax allows a scaffold to be updated by simply editing
the scaffold file. As such scaffolder facilitates a reproducible scaffolding
process and provides a concise overview of how the final genomic scaffold was
constructed. \pb

@@

The scaffold file specifies how sequence regions are joined together; however
the sequences themselves are not stored in the scaffold file and are instead
maintained as a separate FASTA file. Nucleotide sequences in a FASTA-format
file are the standard output for a genome assembler, and these may be contigs,
sets of contigs scaffolded into larger sequences, or both. These nucleotide
sequences stored in the FASTA file are referenced in the scaffold using the
first word from in the FASTA header of the corresponding sequence. Maintaining
the nucleotide sequences in a separate file preserves the unedited sequence
and decouples the data from the specification of how it should be used to
produce the genome sequence. \pb

@@

Scaffolder is software aimed at both bioinformaticians and biologists familiar
with the command line who wish to build a genome scaffold from a set of
contigs. The Scaffolder file format maintains the genome scaffold as a concise
and readable text representation that allows third parties to see how the
genome sequence was scaffolded. This file format also allows a broad overview
of which sequences were included and how they are ordered in the genome
scaffold, something not possible to deduce from a megabase-length string of
nucleotide characters. Scaffolder furthers increases the ease of
reproducibility in genome projects by allowing the scaffold super-sequence to
be reliably reproduced from the same scaffold file. The YAML syntax for
writing the scaffold file is also standardised and simple to manipulate
programmatically. This thereby means the scaffolding process follows the Unix
tenet of: ``If your data structures are good enough, the algorithm to
manipulate them should be trivial.''.

>>>>

  \paragraph*{Methods:} The software outlined here, ``Scaffolder,'' is
  implemented in the Ruby programming language and can be installed via the
  RubyGems software management system. Genome scaffolds are defined using
  YAML - a data format, which is both human and machine-readable. Command line
  binaries and extensive documentation are available.

  \paragraph*{Results:} This software allows a genome build to be defined in
  terms of the constituent sequences using a relatively simple syntax to
  define the scaffold. This syntax further allows  unknown regions to be
  defined, and add additional sequences to fill gaps in the scaffold. Defining
  the genome construction in a file makes the scaffolding process reproducible
  and easier to edit compared with FASTA nucleotide sequence.

@@

The advent of high throughput sequencing methods has led to a renewed interest
in algorithms to solve the problem of genome assembly \cite{pop2008,pop2009}.
Nonetheless, due to the complexity of merging large numbers of overlapping
reads genome assembly software may be unable to produce a complete sequence.
Instead, the algorithm may generate several large assembled regions of
sequence (`contigs') composed of the many individual reads that represent
assembly into the longest possible sequences, given the data. Nonetheless,
these contigs represent a fragmented picture of the genome and require
additional work to join together into a complete sequence. \pb

@@

The software outlined here, ``Scaffolder,'' aims to address these problems of
reproducibility by creating a file syntax and software framework for editing
a genome scaffold. Scaffolder uses a specific file format to define how
contigs are joined, additional sequences are inserted, and for the
specification of unknown regions. This syntax allows a scaffold to be updated
by simply editing the scaffold file. As such, scaffolder facilitates
a reproducible scaffolding process and provides a concise overview of how the
final genomic scaffold was constructed. \pb

@@

The scaffold file specifies how sequence regions are joined together; however,
the sequences themselves are not stored in the scaffold file and are instead
maintained as a separate FASTA file. Nucleotide sequences in a FASTA-format
file are the standard output for a genome assembler, and these may be contigs,
sets of contigs scaffolded into larger sequences, or both. These nucleotide
sequences stored in the FASTA file are referenced in the scaffold using the
first word from in the FASTA header of the corresponding sequence. Maintaining
the nucleotide sequences in a separate file preserves the unedited sequence
and decouples the data from the specification of how it should be used to
produce the genome sequence. \pb

@@

Scaffolder is software aimed at both bioinformaticians and biologists familiar
with the command line who wish to build a genome scaffold from a set of
contigs. The Scaffolder file format maintains the genome scaffold as a concise
and readable text representation that allows third parties to see how the
genome sequence was scaffolded. This file format also allows a broad overview
of which sequences were included and how they are ordered in the genome
scaffold, something not possible to deduce from a megabase-length string of
nucleotide characters. Scaffolder furthers increases the ease of
reproducibility in genome projects by allowing the scaffold super-sequence to
be reliably reproduced from the same scaffold file. The YAML syntax for
writing the scaffold file is also standardised and simple to manipulate
programmatically. This thereby means the scaffolding process follows the Unix
tenet of ``If your data structures are good enough, the algorithm to
manipulate them should be trivial.''
